Where Do Population III Stars Form? The Effects of Radiative Feedback and Self-Shielding on the Host Halo Mass Distribution

### 2019-05-13 paper.tex
Adding shielding calculation, adjustments to abstract, conclusion, title

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Where Do Pop III Stars Form?]{Where Do Population III Stars Form?}

@@

% Abstract of the paper
\begin{abstract}
We perform a cosmological simulation with a comoving volume of 1 Mpc$^{3}$ to study the birthplaces of Population III stars, using the adaptive mesh refinement code \enzo{}. We investigate the distribution of host halo masses and its relationship to the Lyman-Werner background intensity.  In our sample of 697 host halos, we find that 84\% of them have masses below the Machacek et al. (2001) relation because of the inclusion of \hh{} self-shielding.  In our simulation above a redshift of 11.5, the mean halo mass is time-independent and ~10$^{5.9}$ M$_{\odot}$. Afterwards, it steadily rises above the Machacek et al. relation to a mean value of ~10$^{6.6}$ M$_{\odot}$. Most of these halos form multiple Population III stars, with a median number of four, up to a maximum of 16. We also find that a few halos do form stars below the Machacek et al. relation but in a high Lyman-Werner radiation field with values up to ~50 J$_{21}$. Our results suggest that Population III star formation may be less affected by Lyman-Werner radiation feedback than previously thought and that Population III multiple systems are common.
\end{abstract}

@@

Upon numerically calculating the shielding factor of \hh{} for simulated protogalaxies, Wolcott-Green et al. determine that a slight adjustment to the shielding factor from \citet{Draine96} is sufficient to account for inaccuracies at higher temperatures. The implemented shielding factor is then (see Eq. 10 in \citet{Wolcott11}):
\begin{equation} \label{shield}
	\begin{multlined}
	f_{\rm sh}(N_{\rm H2}, T) = \frac{0.965}{(1+x/b_{5})^{1.1}} + \frac{0.035}{(1+x)^{0.5}}  \times \\ \exp [-8.5 \times 10^{-4} (1+x)^{0.5}]
	\end{multlined}
\end{equation}
Here, x $\equiv$ N$_{H_{2}}$/5 $\times$ 10$^{14}$ cm$^{-2}$, b$_{5}$ $\equiv$ b/10$^{5}$ cm s$^{-1}$, and $b$ is the Doppler broadening parameter.

%====================================================================
\subsection{Analytic Calculation of the Shielding Factor for a Static Halo}
%====================================================================
To demonstrate how the shielding factor can affect the column density of \hh{}, we can examine $N_{\rm H2}$ to the core as a function of halo mass for an isothermal halo, with a constant \hh{} fraction. The total number of \hh{} molecules within the halo is given by 
\begin{equation} \label{numberh2}
	n_{\rm H2}(r) = f_{\rm H2} \frac{\rho_{0}R^{2}_{\rm vir}}{\mu m_{\rm H} r^{2}}
\end{equation}
where $\mu = 1.22$ is the mean molecular weight, $\rho_{0} = 200 \rho_{c} / 3$ is the density at the virial radius, and $\rho_{c} = 3 \rm{H}_{0}^{2} / 8 \pi \rm{G} (1+z)^3$ is the critical density. The column density is then equation \ref{numberh2} integrated from the radius of the core to the virial radius, where the core has a radius of $f_{c} R_{\rm vir}$ with $f_{c} = 10^{-2}$: 
\begin{equation} \label{columnh2}
	N_{\rm H2} = f_{\rm H2} R_{\rm vir} \frac{200 \rho_{c}}{3 \mu m_{\rm H}} \left(\frac{1}{f_{c}} - 1 \right)
\end{equation}
For $z = 15$ and $\rm{H}_{0} = 70 \rm{km} \rm{s}^{-1} \rm{Mpc}^{-1}$, ... 
%====================================================================
\section{Results}
%====================================================================
In this section, we will present the distribution of host halo masses of Pop III stars, the relation between the mass distribution and the LW background radiation, and the distribution of various Pop III properties.

@@

In redshift bins of $\Delta z$ = 1, the mean halo mass hosting Pop III stars is determined, and plotted as black dots against the redshift bins in Figure \ref{fig:mean_mass}. The median for each bin is represented by an x, and the 15.9 and 84.1 percentiles are plotted below and above the mean, respectively. The M01 relation (Equation \ref{mthresh}) for our LW background intensity in Eq. \ref{LWbg} is shown as the black solid line. Notably, the mean halo mass falls well below the relation, at $7.7 \times 10^{5}$ M$_{\odot}$ until $z$ = 11.5, at which point, the mean halo mass rises to $3.7 \times 10^{6}$ M$_{\odot}$. The large discrepancy between our data and the mass threshold from M01 is indicative of the \hh{} shielding which is included in our simulation. \hh{} shielding allows for halos to form at much lower masses, and therefore, Pop III stars are forming in these halos at earlier times. In the M01 analysis, \hh{} shielding is neglected in their calculations for a variety of reasons, including the Doppler broadening of LW bands and large column densities of \hh{} only beginning to form at late times. Interestingly, a full radiative transfer simulation, such as ours, does not produce identical results as M01 predicts. Throughout the literature, there appears to be a consensus that \hh{} self-shielding will help smaller mass halos collapse in the presence of a LW background radiation \citep[E.g.][]{Yoshida03, Ricotti01, Glover01, Hartwig15}. As can be seen by the red solid line in \ref{fig:mean_mass}, nearly 100\% of the halos hosting Pop III stars fall below the M01 relation, until around a redshift of 15, at which point, the mean halo mass begins to rise above the relation. The increase in the mean halo mass above the M01 relation is indicative of the increased LW intensity permeating the box due to a galaxy that becomes very active in star formation at late times. This galaxy has a halo mass of $3.6 \times 10^{8} M_{\odot}$, a total stellar mass of $3.1 \times 10^{6} M_{\odot}$, and has a peak Pop III SFR of $2.7 \times 10^{-4} M_{\odot} yr^{-1}$ at $\approx$ 250 Myr and a peak metal enriched SFR of $5.4 \ times 10^{-2}$ at $\approx$ 500 Myr. This galaxy produces a large amount of LW radiation which dominates the box, and drives up the minimum halo mass required for the formation of Pop III stars. This increase in the LW background can also be seen in Figure \ref{fig:JLW_xe_mass}, where by the end of the simulation, the LW background rises to greater than 1 $J_{21}$.

@@

We now inspect the number of Pop III stars and the total mass of Pop III stars per halo. Again, we are only tracking massive stars, not any stars that may form out of fragmented gas with masses < $1 M_{\odot}$. Figure \ref{fig:totnump3_halomass_sidehist} shows the number of Pop III stars per halo for a given halo mass, just after star formation. The histogram of the number of Pop III stars for all masses is projected on the right hand side. We find that a median number of four Pop III stars form per halo, with a maximum of 16 Pop III stars forming per halo. Since we did not restrict the number of Pop III stars that can form in a halo, we find that the conditions are often sufficient for multiple Pop III stars to form in a single halo. Out of the halos forming Pop III stars, only 16\% of them form single Pop III stars, whereas 54\% form between two and five Pop III stars. It should be noted that these halos aren't necessarily forming all of the Pop III stars at once. For example, a halo can form stars again if it din't host a supernova. The number of Pop III stars that form within a halo is indicative of the star formation history of that halo. Figure \ref{fig:totp3mass_halomass_sidehist} shows the total mass of Pop III stars per halo for a given halo mass, where the histogram of the total Pop III mass for all halos is projected on the right hand side. Lines of constant star formation efficiency are overplotted. We find that most of our Pop III stars are forming between 10$^{-4}$ < f$_\star$ < 10$^{-3}$. The mean total mass of Pop III stars for all halos is 195 M$_{\odot}$.

At the end of the simulation, Figure \ref{fig:final_redshift_Np3_mass} shows the distribution of the number of Pop III stars per halo for a given halo mass, for all halos in the simulation that host either a living or a remnant Pop III star. The red line shows the median number of Pop III stars for each halo mass bin. At low masses, the number of Pop III stars per halo sits around 2 Pop III stars, and quickly rises once the halo mass reaches $10^{7} M_{\odot}$. We can also look at the spread in creation times of these Pop III stars per halo, shown in Figure \ref{fig:p3spread_mass}. The halos that have a zero spread in their Pop III creation time, meaning the Pop III stars formed at the same time, lie at $10^{-2}$ Myr. The red line indicates the median spread in creation times for each halo mass bin. There are two groups within this plot. The first group contains halos that form their Pop III stars in a spread of < 10 Myr, with a median spread of about $10^{-1}$ Myr. The second group contains halos that form their Pop III stars in a spread of > 10 Myr, with a median spread of about 100 Myr. The second group represents typically larger mass halos that have acquired older Pop III stars by the end of the simulation through their long merger history. These halos have acquired these Pop III stars from outside halos, accounting for the large spread in Pop III creation time as well as the increased number of Pop III stars, as seen in Figure \ref{fig:final_redshift_Np3_mass}.

@@

\begin{figure*}
	\centering
	\begin{subfigure}{.5\textwidth}
  		\centering
  		\includegraphics[width=\columnwidth]{images/totnump3_halomass_sidehist.pdf}
  		\caption{Total number of Pop III stars in halos hosting new Pop III stars versus halo mass at the instance of star formation.}
  		\label{fig:totnump3_halomass_sidehist}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
  		\centering
  		\includegraphics[width=\columnwidth]{images/final_redshift_Np3_mass.pdf}
  		\caption{Total number of Pop III stars in all halos versus halo mass at the end of the simulation, at $z \approx 9$.}
  		\label{fig:final_redshift_Np3_mass}
	\end{subfigure}
	\caption{The number of Pop III stars per halo versus halo mass at the instance of Pop III star formation and the final redshift. On the left, a Pop III star will form in a halo which contains a median number of four Pop III stars, with some containing as many as 16 Pop III stars. On the right, the red line indicates the median number of Pop III stars in each halo mass bin.}
\end{figure*}

@@

%====================================================================
\section{Conclusions}
%====================================================================
In this work, we presented the analysis of the host halo mass distribution of Pop III stars and how it relates to the LW background radiation. We find that Pop III stars are forming in halos with a mean mass of $10^{5.9} M_\odot$ until $z = 11.5$, which falls well below the M01 relation due to \hh{} self-shielding. The mean halo mass then rises above the M01 to a mean value of $10^{6.6} M_\odot$ due to the domination of LW radiation throughout our simulation box produced by metal-enriched stars. As we compare with the literature, we appear to have much lower halo masses than those that do not include \hh{} self-shielding, so it appears that this is an important process that is often ignored in calculations. We also see that there doesn't appear to be a correlation between the average LW intensity with a halo and the halo mass. Therefore, the LW intensity may not be as indicative of which halos may form Pop III stars as previously thought. Pop III stars are often assumed to form in isolation, but we are finding that most of our halos are forming multiple Pop III stars, with a median number of four Pop III stars in a single halo, up to a maximum of 16 Pop III stars.

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Where Do Pop III Stars Form?]{Where Do Population III Stars Form? The Effects of Radiative Feedback and Self-Shielding on the Host Halo Mass Distribution}

@@

% Abstract of the paper
\begin{abstract}
The formation of Population III (Pop III) stars is a critical step in the evolution of the early universe. To understand how these stars affected their metal-enriched descendants, the details of how, why and where Pop III formation takes place is crucial to understand. One of the processes that is assumed to greatly affect the formation of Pop III stars is the presence of a Lyman-Werner (LW) radiation background, which can destroy \hh{}, a necessary ingredient for the cooling and creation of Pop III stars. \hh{} self-shielding can alleviate the effect the LW background has on the \hh{} within halos, and is a process which is often neglected throughout the literature. In this work, we perform a cosmological simulation with a comoving volume of 1 Mpc$^{3}$ to study the birthplaces of Population III stars, using the adaptive mesh refinement code \enzo{}. We investigate the distribution of host halo masses and its relationship to the Lyman-Werner background intensity. In our simulation above a redshift of 11.5, the mean halo mass is time-independent and $7.7 \times 10^{5} M_{\odot}$. Afterwards, it steadily rises to a mean value of $3.7 \times 10^{6} M_{\odot}$, due to the increased LW background coming from a massive galaxy actively forming stars. Compared to the literature, our halos form Pop III stars at much lower masses due tot eh inclusion of \hh{} self-shielding. Most halos form multiple Population III stars, with a median number of four, up to a maximum of 16, at the instance of Pop III formation. We also find that a few halos do form stars in a high Lyman-Werner radiation field with values up to ~50 J$_{21}$. Our results suggest that Population III star formation may be less affected by Lyman-Werner radiation feedback than previously thought and that Population III multiple systems are common. 
\end{abstract}{}

@@

Upon numerically calculating the shielding factor of \hh{} for simulated protogalaxies, Wolcott-Green et al. determine that a slight adjustment to the shielding factor from \citet{Draine96} is sufficient to account for inaccuracies at higher temperatures. The implemented shielding factor is then (see Eq. 10 in \citet{Wolcott11}):
\begin{equation} \label{shield}
	\begin{multlined}
	f_{\rm sh}(N_{\rm H2}, T) = \frac{0.965}{(1+x/b_{5})^{1.1}} + \frac{0.035}{(1+x)^{0.5}}  \times \\ \exp [-8.5 \times 10^{-4} (1+x)^{0.5}]
	\end{multlined}
\end{equation}
Here, x $\equiv$ N$_{H_{2}}$/5 $\times$ 10$^{14}$ cm$^{-2}$, b$_{5}$ $\equiv$ b/10$^{5}$ cm s$^{-1}$, and $b$ is the Doppler broadening parameter. This shielding factor acts as a multiplier to the \hh{} dissociation rate, where if the shielding factor equals one, there is no shielding, and zero if there is maximum shielding.

%====================================================================
\subsection{Analytic Calculation of the Shielding Factor for a Static Halo}
%====================================================================
To demonstrate how the shielding factor changes as a function of halo mass and redshift, we can examine $N_{\rm H2}$ to the core as a function of halo mass for an isothermal halo, with a constant \hh{} fraction, and use Equation \ref{shield} to calculate the shielding factor. The total number of \hh{} molecules within the halo is given by 
\begin{equation} \label{numberh2}
	n_{\rm H2}(r) = f_{\rm H2} \frac{\rho_{0}R^{2}_{\rm vir}}{\mu m_{\rm H} r^{2}}
\end{equation}
where $\mu = 1.22$ is the mean molecular weight, $\rho_{0} = 200 \rho_{c} / 3$ is the density at the virial radius, and $\rho_{c} = 3 \rm{H}_{0}^{2} / 8 \pi \rm{G} (1+z)^3$ is the critical density. The column density is then equation \ref{numberh2} integrated from the radius of the core to the virial radius, where the core has a radius of $f_{c} R_{\rm vir}$: 
\begin{equation} \label{columnh2}
	N_{\rm H2} = f_{\rm H2} R_{\rm vir} \frac{200 \rho_{c}}{3 \mu m_{\rm H}} \left(\frac{1}{f_{c}} - 1 \right)
\end{equation}

For $\rm{H}_{0} = 70 \rm{km} \rm{s}^{-1} \rm{Mpc}^{-1}$, a typical \hh{} fraction of $f_{\rm H2} = 10^{-4}$, and assuming the Doppler broadening parameter, $b$, is equal to the circular velocity of the halo, Equation \ref{columnh2} and Equation \ref{shield} will give the shielding factor for a given core radius, $f_{c}$, and redshift. Figure \ref{fig:shield_mass} shows the shielding factor as a function of halo mass, for redshifts $z = 9$, $15$, and $25$ in red, blue and black, respectively. The solid lines indicate a core radius of $f_{c} = 10^{-2}$ and the dashed lines a core radius of $f_{c} = 10^{-4}$. At each redshift, all halo masses, and each core radius, we see the shielding factor is significantly smaller than one, implying that \hh{} shielding for these halos is almost at a maximum. The shielding factor will then significantly decrease the dissociation rate of \hh{}, allowing for the presence of \hh{} within a LW background. Smaller mass halos can then form Pop III stars since \hh{} will stick around in the halo for a longer amount of time.

\begin{figure}
	\includegraphics[width=\columnwidth]{images/shield_mass.pdf}
    \caption{Shielding factor as a function of halo mass. Solid lines indicate a core radius of $f_{c} = 10^{-2}$ and the dashed lines indicate a core radius of $f_{c} = 10^{-4}$. For these parameters, and assuming the Doppler broadening parameter equals the circular velocity, the shielding factor is almost zero, implying maximum shielding.}
    \label{fig:shield_mass}
\end{figure}
%====================================================================
\section{Results}
%====================================================================
In this section, we will present the distribution of host halo masses of Pop III stars, the relation between the mass distribution and the LW background radiation, and the distribution of various Pop III properties.

@@

In redshift bins of $\Delta z$ = 1, the mean halo mass hosting Pop III stars is determined, and plotted as black dots against the redshift bins in Figure \ref{fig:mean_mass}. The median for each bin is represented by an x, and the 15.9 and 84.1 percentiles are plotted below and above the mean, respectively. The M01 relation (Equation \ref{mthresh}) for our LW background intensity in Eq. \ref{LWbg} is shown as the black solid line. Notably, the mean halo mass falls well below the relation, at $7.7 \times 10^{5}$ M$_{\odot}$ until $\approx$ 380 Myr, at which point, the mean halo mass rises to $3.7 \times 10^{6}$ M$_{\odot}$. The large discrepancy between our data and the mass threshold from M01 is indicative of the \hh{} shielding which is included in our simulation. \hh{} shielding allows for halos to form at much lower masses, and therefore, Pop III stars are forming in these halos at earlier times. In the M01 analysis, \hh{} shielding is neglected in their calculations for a variety of reasons, including the Doppler broadening of LW bands and large column densities of \hh{} only beginning to form at late times. Interestingly, a full radiative transfer simulation, such as ours, does not produce identical results as M01 predicts. Throughout the literature, there appears to be a consensus that \hh{} self-shielding will help smaller mass halos collapse in the presence of a LW background radiation \citep[E.g.][]{Yoshida03, Ricotti01, Glover01, Hartwig15}. As can be seen by the red solid line in \ref{fig:mean_mass}, nearly 100\% of the halos hosting Pop III stars fall below the M01 relation, until around a redshift of 15, at which point, the mean halo mass begins to rise above the relation. The increase in the mean halo mass above the M01 relation is indicative of the increased LW intensity permeating the box due to a galaxy that becomes very active in star formation at late times. This galaxy has a halo mass of $3.6 \times 10^{8} M_{\odot}$, a total stellar mass of $3.1 \times 10^{6} M_{\odot}$, and has a peak Pop III SFR of $2.7 \times 10^{-4} M_{\odot} yr^{-1}$ at $\approx$ 250 Myr and a peak metal enriched SFR of $5.4 \times 10^{-2}$ at $\approx$ 500 Myr. This galaxy produces a large amount of LW radiation which dominates the box, and drives up the minimum halo mass required for the formation of Pop III stars. This increase in the LW background can also be seen in Figure \ref{fig:JLW_xe_mass}, where by the end of the simulation, the LW background rises to greater than 1 $J_{21}$.

@@

We now inspect the number of Pop III stars and the total mass of Pop III stars per halo. Again, we are only tracking massive stars, not any stars that may form out of fragmented gas with masses < $1 M_{\odot}$. The left panel of Figure \ref{fig:num_p3} shows the number of Pop III stars per halo for a given halo mass, just after star formation. The histogram of the number of Pop III stars for all masses is projected on the right hand side. We find that a median number of four Pop III stars form per halo, with a maximum of 16 Pop III stars forming per halo. Since we did not restrict the number of Pop III stars that can form in a halo, we find that the conditions are often sufficient for multiple Pop III stars to form in a single halo. Out of the halos forming Pop III stars, only 16\% of them form single Pop III stars, whereas 54\% form between two and five Pop III stars. It should be noted that these halos aren't necessarily forming all of the Pop III stars at once. For example, a halo can form stars again if it din't host a supernova. The number of Pop III stars that form within a halo is indicative of the star formation history of that halo. Figure \ref{fig:totp3mass_halomass_sidehist} shows the total mass of Pop III stars per halo for a given halo mass, where the histogram of the total Pop III mass for all halos is projected on the right hand side. Lines of constant star formation efficiency are overplotted. We find that most of our Pop III stars are forming between 10$^{-4}$ < f$_\star$ < 10$^{-3}$. The mean total mass of Pop III stars for all halos is 195 M$_{\odot}$.

At the end of the simulation, the right panel of Figure \ref{fig:num_p3} shows the distribution of the number of Pop III stars per halo for a given halo mass, for all halos in the simulation that host either a living or a remnant Pop III star. The red line shows the median number of Pop III stars for each halo mass bin. At low masses, the number of Pop III stars per halo sits around 2 Pop III stars, and quickly rises once the halo mass reaches $10^{7} M_{\odot}$. We can also look at the spread in creation times of these Pop III stars per halo, shown in Figure \ref{fig:p3spread_mass}. The halos that have a zero spread in their Pop III creation time, meaning the Pop III stars formed at the same time, lie at $10^{-2}$ Myr. The red line indicates the median spread in creation times for each halo mass bin. There are two groups within this plot. The first group contains halos that form their Pop III stars in a spread of < 10 Myr, with a median spread of about $10^{-1}$ Myr. The second group contains halos that form their Pop III stars in a spread of > 10 Myr, with a median spread of about 100 Myr. The second group represents typically larger mass halos that have acquired older Pop III stars by the end of the simulation through their long merger history. These halos have acquired these Pop III stars from outside halos, accounting for the large spread in Pop III creation time as well as the increased number of Pop III stars, as seen in the right panel of Figure \ref{fig:num_p3}.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.48\textwidth]{images/totnump3_halomass_sidehist.pdf}
  \hfill
  \includegraphics[width=0.48\textwidth]{images/final_redshift_Np3_mass.pdf}
  \caption{The number of Pop III stars per halo versus halo mass at the instance of Pop III star formation (left) and the final redshift (right). At the instance of formation, Pop III stars will form in a halo which contains a median number of four Pop III stars, with some containing as many as 16 Pop III stars. On the right, the red line indicates the median number of Pop III stars in each halo mass bin. By the end of the simulation, high mass halos contain a large number of Pop III stars due to their long merger history with smaller halos hosting Pop III stars.}
  \label{fig:num_p3}
\end{figure*}

@@

%====================================================================
\section{Conclusions}
%====================================================================
In this work, we presented the analysis of the host halo mass distribution of Pop III stars and how it relates to the LW background radiation. We find that Pop III stars are forming in halos with a mean mass of $7.7 \times 10^{5}$ M$_{\odot}$, which falls well below the M01 relation due to \hh{} self-shielding, until $\approx$ 380 Myr, at which point the mean halo mass rises above the M01 relation to a mean value of $3.7 \times 10^{6}$ M$_{\odot}$ due to the domination of LW radiation produced by metal-enriched stars in the most massive halo in the simulation box. As we compare with the literature, we see much lower halo masses hosting Pop III stars due to the inclusion of \hh{} shielding. It appears that this is an important process that is often ignored in calculations. We also see that there doesn't appear to be a correlation between the average LW intensity within a halo and the halo mass. Therefore, the LW intensity may not be as indicative of which halos may form Pop III stars as previously thought. Pop III stars are often assumed to form in isolation, but we find that most of our halos are forming multiple Pop III stars, with a median number of four Pop III stars in a single halo, up to a maximum of 16 Pop III stars. By the end of the simulation, halos with masses > $10^{7} M_{\odot}$ acquire multiple Pop III stars due to mergers with smaller mass halos. This results in these high mass halos containing multiple Pop III stars, both young and old.

Our results provide another piece of the Pop III puzzle. With the inclusion of \hh{} self-shielding, we find Pop III stars forming in smaller mass halos than previously predicted and not necessarily in isolation. The assumptions often made about the formation of Pop III stars appear to break down when important processes are taken into account, and restrictions about their formation are lifted. Further work needs to be done to determine why some halos form Pop III stars and others don't, to provide further insight about the formation process of these stars and thus how they will affect their surrounding galactic environment.
